---
title: "Project"
output: pdf_document
date: "2023-04-05"
---

# LOAN RISK PREDICTION MODEL USING DECISION TREE

## DATA LOADING
```{r}
data=read.csv("credit_risk_dataset.csv")
```

```{r}
head(data)
```

## DATA PREPROCESSING
```{r}
#Checking for any missing values in the dataset.
sum(is.na(data))
```
There are missing values in the dataset. We need to remove the missing values from the dataset.

```{r}
#Removing rows with missing values
data=na.omit(data)
sum(is.na(data))
```

All the missing values have been removed from the dataset.

```{r}
#Checking the datatype of each column in R
str(data)
```

```{r}
#convert the categorical variables to factors
data$cb_person_default_on_file=as.factor(data$cb_person_default_on_file)
data$person_home_ownership=as.factor(data$person_home_ownership)
data$loan_intent=as.factor(data$loan_intent)
data$loan_grade=as.factor(data$loan_grade)
```

All the datatypes are correct, So no need of changing. The data pre processing is done.

## DATA VISUALISATION


### PLOTTING A SCATTERPLOT BETWEEN THE INCOME AND LOAN AMOUNT
```{r}
library(ggplot2)
pt=ggplot(data=data)+geom_point(aes(y = data$person_income, x = data$loan_amnt))+labs(y = "Person Income", x = "Loan Amount", title = "Scatterplot of Person Income vs Loan Amount")
print(pt)
```

```{r}
pt2=ggplot(data=data)+geom_boxplot(aes(x = data$loan_grade, y = data$loan_int_rate))+labs(x = "Loan Grade", y = "Loan Interest Rate", title = "Boxplot of Loan Grade vs Loan Interest Rate")
plot(pt2)
```
We can see that the Interest Rates are considerably increasing with the level of Loan Grades.

## DATA SPLITTING

We have a large dataset of nearly 30,000 observations. So we can use more data to train. The preferable ratio will be 90% to training data and 10% to testing data.

```{r}
library(caret)
train_index <- createDataPartition(data$loan_status, p = 0.9, list = FALSE)
train_data <- data[train_index, ] 
test_data <- data[-train_index, ]
```


## BUILDING THE FIRST MODEL

```{r}
library(rpart)#required library
library(rpart.plot)
variable=data$cb_person_default_on_file
```


```{r}
model1= rpart(cb_person_default_on_file ~ ., data = train_data)
```
```{r}
rpart.plot(model1)
```

## MODEL-1 PREDICTION AND EVALUATION

```{r}
predict1 =predict(model1, test_data, type = "class")
```

```{r}
# Compute confusion matrix and accuracy for the Decision Tree model
confmat1 = table(test_data$cb_person_default_on_file, predict1)
acc1 = sum(diag(confmat1)) / sum(confmat1)
print(paste("Accuracy:", round(acc1, 3)*100,"%"))
```
## MODEL-2 BUILDING - NAIVE BAYES 

```{r}
library(e1071)
model2 = naiveBayes(cb_person_default_on_file ~ ., data = train_data)
pred2 = predict(model2, test_data)

```

## MODEL-2 EVALUATION
```{r}
# Compute confusion matrix for Naive Bayes Model
confmat2 <- table(test_data$cb_person_default_on_file,pred2)
print(confmat2)

# Compute accuracy for the model
acc2 <- sum(diag(confmat2)) / sum(confmat2)
print(paste("Accuracy:", round(acc2, 3)*100,"%"))


```

